Comparing Cross Validation Methods for Penalized Cox Regression

### 2017-10-03 cross-validation.tex
Add graphs for stability
correct index and change figures

In the Cox model \citep{Cox1975}, the hazard function for an individual observation j is given by \begin{equation}h_{j}(t) = h_{0}(t) \text{exp}( X_{j}^{T} \beta)\end{equation} where $h_{0}$ is the baseline hazard and $e^{X_{j}^{T} \beta}$ is the relative risk. The estimation of the coefficients $\beta$s are obtained by maximizing the partial likelihood \begin{equation}L(\beta) = \prod_{j=1}^{n} \frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}\end{equation} As a semi-parametric model, Cox regression only gives estimations of the $\beta$ coefficients, without estimating the baseline hazard. Hence the interpretation of the Cox model is only valid in a relative sense.

Penalized Cox Regression is an extension of the regular cox regression model. Estimates of the $\beta$ coefficients are obtained by minimizing the objective function \begin{equation}  Q(\beta |X, y) = - \frac{1}{n}  \sum_{j=1}^{n} \text{log} \left \{\frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}) \right \} + P_{\lambda}(\beta). \end{equation}, instead of optimizing the partial likelihood. The first part of the objective function is based on the partial likelihood. The second part is a penalty term that depends on some regularization parameters. In this paper, only the LASSO penalty $P_{\lambda}(\beta) = \lambda (\sum_{j} |\beta_{j}|)$ was considered. The penalized cox regression can be used for analyzing high dimensional data, where the number of covariates p $\gg$ n. A common application would be to predict the overall survival time of a particular cancer by certain gene expressions.

@@

Besides cross validating over the partial likelihood, an alternative approach is to cross validate over the linear predictors. The data would still be split into K folds. Suppose ith fold is left out and the other K-1 folds are used as the training set to get the estimates $\hat{\beta}_{-i}$. Then the cross-validated linear predictors would be calculated based on the observations in the test set:  \begin{equation}\hat{\eta}_{-i} = X_{i}\hat{\beta}_{-i}.\end{equation} After repeating this for all K folds, a whole set of linear predictors  $\hat{\eta}_{-} = ( \hat{\eta}_{-1},  \hat{\eta}_{-2} , ...  \hat{\eta}_{-K})$ can be obtained. A partial likelihood can be built over this set of linear predictors: \begin{equation} L(\hat{\eta}_{-}) = \prod_{j=1}^{n} \frac{exp (\hat{\eta}_{-j})}{\sum_{ k \in R(t_{j})}exp (\hat{\eta}_{-k})}\end{equation}.

@@

\par The original idea of calcualting the Martingale and Deviance residuals were integrated into cross validation. Based on the observations in the training set, a baseline hazard $\hat{\Lambda}_{0}$ can be estimated via Kalbfleisch and Prentice's method \citep{Kalbfleisch2011}. Then for each observation j in the test set, the Martingale Residual can be calculated: \begin{equation}\hat{M_{j}} = d_{j} -\hat{\Lambda}_{0}(t_{j})e^{\hat{\eta}_{j}},\end{equation} where $d_{j}$ is the status of the jth observation and $t_{j}$ is the time component of the jth observation. The deviance residual can be calcualted from the Martingale Residuals: \begin{equation} d_{i} = sgn(\hat{M}_{i})(-2(\hat{M}_{i} + \delta_{i}log(\delta_{i} - \hat{M}_{i})))^{1/2}\end{equation}

After repeating this procedure for all K folds, we would obtain a set of deviance residuals for all observations. They are be pulled together by sum of squares as the cross validated error:
\begin{equation}CVE = \sum_{j}\hat{d}_{j}^2\end{equation}.

@@

@@

\section{Discussion}

>>>>

In the Cox model \citep{Cox1975}, the hazard function for an individual observation i is given by \begin{equation}h_{i}(t) = h_{0}(t) \text{exp}( X_{i}^{T} \beta)\end{equation} where $h_{0}$ is the baseline hazard and $e^{X_{i}^{T} \beta}$ is the relative risk. Suppose there are n individual observations and m observed failure times, the estimation of the coefficients $\beta$s are obtained by maximizing the partial likelihood \begin{equation}L(\beta) = \prod_{j=1}^{m} \frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}\end{equation}, where j indexes the observed failure times and $R(t_{j})$ denotes the set of observations at risk at time $t_{j}$. As a semi-parametric model, Cox regression only gives estimations of the $\beta$ coefficients, without estimating the baseline hazard. Hence the interpretation of the Cox model is only valid in a relative sense.

Penalized Cox Regression is an extension of the regular cox regression model. Estimates of the $\beta$ coefficients are obtained by minimizing the objective function \begin{equation}  Q(\beta |X, y) = - \frac{1}{n}  \sum_{j=1}^{m} \text{log} \left \{\frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}) \right \} + P_{\lambda}(\beta). \end{equation} \citep{Simon2011}. The first part of the objective function is based on the partial likelihood. The second part is a penalty term that depends on some regularization parameters. In this paper, only the LASSO penalty $P_{\lambda}(\beta) = \lambda (\sum_{j} |\beta_{j}|)$ was considered. The penalized cox regression can be used for analyzing high dimensional data, where the number of covariates p $\gg$ n. A common application would be to predict the overall survival time of a particular cancer by certain gene expressions.

@@

Besides cross validating over the partial likelihood, an alternative approach is to cross validate over the linear predictors. The data would still be split into K folds. Suppose ith fold is left out and the other K-1 folds are used as the training set to get the estimates $\hat{\beta}_{-i}$. Then the cross-validated linear predictors would be calculated based on the observations in the test set:  \begin{equation}\hat{\eta}_{-i} = X_{i}\hat{\beta}_{-i}.\end{equation} After repeating this for all K folds, a whole set of linear predictors  $\hat{\eta}_{-} = ( \hat{\eta}_{-1},  \hat{\eta}_{-2} , ...  \hat{\eta}_{-K})$ can be obtained. A partial likelihood can be built over this set of linear predictors: \begin{equation} L(\hat{\eta}_{-}) = \prod_{j=1}^{m} \frac{exp (\hat{\eta}_{-j})}{\sum_{ k \in R(t_{j})}exp (\hat{\eta}_{-k})}\end{equation}.

@@

\par The original idea of calcualting the Martingale and Deviance residuals were integrated into cross validation. Based on the observations in the training set, a baseline hazard $\hat{\Lambda}_{0}$ can be estimated via Kalbfleisch and Prentice's method \citep{Kalbfleisch2011}. Then for each observation in the test set, the Martingale Residual can be calculated: \begin{equation}\hat{M_{i}} = \delta_{i} -\hat{\Lambda}_{0}(t_{i})e^{\hat{\eta}_{i}},\end{equation} where $\delta_{j}$ is the status of the jth observation and $t_{j}$ is the time component of the jth observation. The deviance residual can be calcualted from the Martingale Residuals: \begin{equation} d_{i} = sgn(\hat{M}_{i})(-2(\hat{M}_{i} + \delta_{i}log(\delta_{i} - \hat{M}_{i})))^{1/2}\end{equation}

After repeating this procedure for all K folds, we would obtain a set of deviance residuals for all observations. They are be pulled together by sum of squares as the cross validated error:
\begin{equation}CVE = \sum_{i}\hat{d}_{i}^2\end{equation}.

@@

    \subsubsection {Stability}
   While the ungrouped cross-validated likelihood is an intuitive way to carry out cross-validation, it is quite unstable when it runs into cases where the number of observed events in some fold is really small. As is illustrated in Figure 5, cases of undefined cross-validation error would increase when censoring increases or when number of folds increases.

\begin{figure}
    \centering
		\includegraphics[height= 8cm ]{./figures/stability2.png}
    \caption{Comparing Stability}
\end{figure}

@@

\section{Discussion}
	\begin{itemize}
		\item 	
		\item
	\end{itemize}
